Lifting CDCL to Numerical Domains

### 2016-10-09 algorithm.tex
acdcl algorithm

\section{Abstract Conflict Driven Clause Learning}
\rmcmt{safety as satisfiability}

@@

@@

@@

\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}

@@

\assign result=\deduce(P)\;
\uIf{$result == \perp$} {

@@

\uElse
  \uIf{is\_gamma\_complete()} {
    \return \unsafe \;

@@

  \assign decision\_variable = decision\_heuristics($P$)\;
  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
  \uIf{decision\_variable == NULL}
    return \unknown\; 

@@

  \uElse
    \assign result=\deduce(P)\;
    \uIf{result == UNKNOWN}
      \uIf{is\_gamma\_complete()} {
        \return \unsafe \;
      \uElse 
        \continue\;
    \uElse
      \Do{$result == \perp$} {

@@

        \assign result=\deduce(P)\;

@@

\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}

@@

>>>>

\section{Abstract Conflict Driven Clause Learning}
In a typical abstract interpretation based galois-connection settings over 
an over-approximate domain, every concrete elements have a unique over-approximate 
representation in the abstract.  Likewise, every concrete transformer is
over-approximated by a unique abstract transformer.  We now define a abstract 
deduction transformer.

\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
domain $A$ is a sound approximation of a concrete model transformer
$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
$m \in A$ is a meet irreducible.   
\end{definition}

Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
\rangle$.  An abstract deduction transformer is typically computed in the form 
of strongest post-condition or a weakest pre-condition of a formula in the 
abstract domain.

A meet decomposition of the outcome of abstract deduction transformer 
is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
precisely complementable.

\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
\end{definition}

For a program with $N$ variables, let $L$ be the total number of 
meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
of all octagonal inequalities in the set $L$.
%\rmcmt{safety as satisfiability}

@@

\SetKwData{conflict}{CONFLICT}

@@

\SetKwFor{Loop}{Loop}{}{}

@@

\Input{Set $S$ of abstract deduction transformer, 
$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract domain $A$}

@@

$v \leftarrow \top$ \;
$\mathcal{T} \leftarrow \epsilon$ \;
$\mathcal{R} \leftarrow \emptyset$ \;
$result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;
\uIf{$result$ == \conflict} {

@@

@@

  \uIf{$v$ is $\gamma$-complete} {
    \return \unsafe;

@@

  $d \leftarrow decide(v)$ \;
  $v \leftarrow v \meet d$ \; 
  $\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
  $\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
  $result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
  \Do{$result == \conflict$} {
    learn() \;

@@

    result $\leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;

@@

\caption{Abstract Conflict Driven Clause Learning $\{ACDCL(S)\}$ \label{Alg:acdcl}}

@@

A decision is made by calling a procedure $decide(v)$ which 
returns a meet irreducible $d$.  If $d$ refines the 
current abstract value $v$, then the worklist is updated with 
all transformers that are dependant on the current decision 
$d$ along with the update of the corresponding propagation 
and reason trail, $\mathcal{T}$ and $\mathcal{R}$.

### 2016-10-09 deductions.tex
acdcl algorithm

\section{Deduction}
In a typical abstract interpretation based galois-connection settings over 
an over-approximate domain, every concrete elements have a unique over-approximate 
representation in the abstract.  Likewise, every concrete transformer is
over-approximated by a unique abstract transformer.  We now define a abstract 
deduction transformer.

\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
domain $A$ is a sound approximation of a concrete model transformer
$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
$m \in A$ is a meet irreducible.   
\end{definition}

Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
\rangle$.  An abstract deduction transformer is typically computed in the form 
of strongest post-condition or a weakest pre-condition of a formula in the 
abstract domain.

A meet decomposition of the outcome of abstract deduction transformer 
is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
precisely complementable.

\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
\end{definition}

For a program with $N$ variables, let $L$ be the total number of 
meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
of all octagonal inequalities in the set $L$.
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & SSA & Octagon \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
int main() {
 unsigned x, y;
 __CPROVER_assume(x==y);
 x++;
 assert(x==y+1);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
$\begin{array}{l@{\,\,}c@{\,\,}l}
SSA &\iff& ((guard0 == TRUE) \land \\
    &    & (cond == (x == y)) \land \\
    &    & (guard1 == (cond \&\& guard0)) \land \\
    &    & (x' == 1u + x) \land \\
    &    & (x' == 1u + y || !guard1))
\end{array}$
\end{minipage}
&
\begin{minipage}{3.75cm}
$\begin{array}{l@{\,\,}c@{\,\,}l}
C &\iff& ((x' > 1) \land (-x'-y < -2) \land \\
  &    & (-x-x' < -2) \land (y-x' < 0) \land \\                                                                
  &    & (x-x' < 0) \land (y > 0) \land \\
  &    & (x > 0) \land (-x'-y < 0) \land \\
  &    & (x+y > 1) \land (y-x < 1) \land \\
  &    & (x'-y < 2) \land (x-y < 1) \land \\
  &    & (x+y > 0) \land (x+x' > 0) \land \\
  &    & (x'-x < 2))
\end{array}$
\end{minipage}
\\
\hline
\end{tabular}
\caption{C Program, corresponding SSA and Octagonal Inequalities}
\label{swssa}
\end{figure}
%
\subsection{Abstract Model Search}
Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
In propositional case, a model search involves two steps -- 
{\em deductions} using unit rule refines current partial assignments 
and {\em decisions} to heuristically guess a value for an unassigned 
literal.  A unit rule overapproximates a model transformer and deduction 
computes a greatest fixed point over the partial assignments
domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
a greatest fixed point over meet irreducible deduction transformer in $S$.

A {\em worklist} is initialized following a heuristics choice for iteration 
strategies. A {\em forward} iteration strategy initializes the worklist with 
transformers that has constants in the right-hand side. Whereas, a {\em backward} 
iteration strategy starts with a set of assertions. On the other hand, a 
{\em chaotic} iteration strategy initializes the worklist with set of all 
transformers.  For every meet irreducible deduction transformer $s \in S$ 
passed to the abstract domain, if the domain infers a new meet irreducible 
$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
is inferred by the domain, the algorithm terminates with UNSAT.  However,
when a fixed-point is reached, the abstract value $v$ is checked for set of
models.  This is done by checking whether the set of assignments in $v$ is 
$\gamma$-complete~\cite{gamma}.  If $v$ is $\gamma$-complete, the abstract model 
search terminates with SAT.  Otherwise, a decision is made by calling a 
procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
current abstract value $v$, then the corresponding propagation and reason trail,
$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure terminates with UNKNOWN.      
%    
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{deduce}{deduce}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwData{sat}{SAT}
\SetKwData{unsat}{UNSAT}
\SetKwData{unknown}{UNKNOWN}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
abstract value $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
\Output{The status (\sat or \unsat or \unknown)}
worklist $\leftarrow$ initialize\_worklist($S$) \;
\Loop{}
{
// Greatest fixed point \;
\While{$!worklist.empty()$} 
{
  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
  // check if the domain deduce BOTTOM \;
  \uIf{$meet\_irrd == \bot$} {
    $\mathcal{R}[\bot] \leftarrow s$ \;
    // empty the worklist \;
    worklist.delete() \;
    \return \sat \;
  }
  \uElse
  {
    // add deduction to trail \;
    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
    // update the worklist \;
    worklist.update(v, s) \; 
  }
}
// check for counterexample\;
\uIf{$v$ is $\gamma$-complete} {
  \return \unsat;
}
$d \leftarrow decide(v)$ \;
\uIf{$v \sqsubseteq d$} {
\return \unknown \;
}
// add decision to trail\;
$\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
$v \leftarrow v \meet d$ \; 
$\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
}
\end{small}
\caption{Abstract Model Search Procedure\label{Alg:ms}}
\end{algorithm2e}
%

>>>>

\section{Abstract Model Search}
%
\begin{figure}[t]
\scriptsize
\begin{tabular}{l|l|l}
\hline
C program & SSA & Octagon \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
int main() {
 unsigned x, y;
 __CPROVER_assume(x==y);
 x++;
 assert(x==y+1);
}
\end{lstlisting}
&
\begin{minipage}{4.40cm}
$\begin{array}{l@{\,\,}c@{\,\,}l}
SSA &\iff& ((guard0 == TRUE) \land \\
    &    & (cond == (x == y)) \land \\
    &    & (guard1 == (cond \&\& guard0)) \land \\
    &    & (x' == 1u + x) \land \\
    &    & (x' == 1u + y || !guard1))
\end{array}$
\end{minipage}
&
\begin{minipage}{3.75cm}
$\begin{array}{l@{\,\,}c@{\,\,}l}
C &\iff& ((x' > 1) \land (-x'-y < -2) \land \\
  &    & (-x-x' < -2) \land (y-x' < 0) \land \\                                                                
  &    & (x-x' < 0) \land (y > 0) \land \\
  &    & (x > 0) \land (-x'-y < 0) \land \\
  &    & (x+y > 1) \land (y-x < 1) \land \\
  &    & (x'-y < 2) \land (x-y < 1) \land \\
  &    & (x+y > 0) \land (x+x' > 0) \land \\
  &    & (x'-x < 2))
\end{array}$
\end{minipage}
\\
\hline
\end{tabular}
\caption{C Program, corresponding SSA and Octagonal Inequalities}
\label{swssa}
\end{figure}
%
%    
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{deduce}{deduce}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwData{sat}{SAT}
\SetKwData{conflict}{CONFLICT}
\SetKwData{unsat}{UNSAT}
\SetKwData{unknown}{UNKNOWN}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKwFor{Loop}{Loop}{}{}
\SetKw{KwNot}{not}
\begin{small}
\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
abstract valuation $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
\Output{The status (\sat or \conflict or \unknown)}
worklist $\leftarrow$ initialize\_worklist($S$) \;
\Loop{}
{
// Greatest fixed point \;
\While{$!worklist.empty()$} 
{
  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
  \uIf{$meet\_irrd == \bot$} {
    $\mathcal{R}[\bot] \leftarrow s$ \;
    worklist.delete() \;
    \return \conflict \;
  }
  \uElse
  {
    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
    worklist.update(v, s) \; 
  }
}
\uIf{$v$ is $\gamma$-complete} {
  \return \sat;
}
\uElse {
 \return \unknown \;
}
}
\end{small}
\caption{Abstract Model Search $\{deduce(S,v,\mathcal{T},\mathcal{R})\}$ \label{Alg:ms}}
\end{algorithm2e}
%  
A model search procedure in a SAT solver involves two steps -- {\em deductions} 
using unit rule refines current partial assignments and 
{\em decisions} to heuristically guess a value for an unassigned 
literal.  A unit rule overapproximates a model transformer and deduction 
computes a greatest fixed point over the partial assignments
domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
that computes a greatest fixed point over meet irreducible deduction 
transformer in $S$.  
%
\subsection{Algorithm for Deduction Phase}
%
Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.
A {\em worklist} is initialized following a heuristics choice for iteration 
strategies. A {\em forward} iteration strategy initializes the worklist with 
transformers that has constants in the right-hand side. Whereas, a {\em backward} 
iteration strategy starts with a set of assertions. On the other hand, a 
{\em chaotic} iteration strategy initializes the worklist with set of all 
transformers.  For every meet irreducible deduction transformer $s \in S$ 
passed to the abstract domain, if the domain infers a new meet irreducible 
$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
is inferred by the domain, the algorithm terminates with CONFLICT.  However,
when a fixed-point is reached, the abstract valuation $v$ is checked for set of
models.  This is determined by checking whether the set of assignments in $v$ is 
$\gamma$-complete~\cite{gamma}.  This can be achieved by checking if there is a 
concrete solution $c$ such that the inequality $c(x) \subseteq \gamma \circ
v(x)$ holds for every meet irreducible $x$.  If $v$ is $\gamma$-complete, that is, the 
abstract valuation $v$ is abstractly satisfying, the abstract model 
search terminates with SAT.  Otherwise, the algorithm returns UNKNOWN and the 
model search procedure makes a new decision.

### 2016-10-09 domains.tex
acdcl algorithm

APRON library is specialized for the abstract interpretation 
framework which does not require complementation 
operator.  To this end, we implement our own template-based 
octagon domain which supports complementation operator.  
For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
The complementation of the above octagon can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

>>>>

APRON library is specialized for the abstract interpretation 
framework which does not require complementation property.  
To this end, we implement our own template-based 
interval, octagon and equality domains which supports 
complementation operator.  For example, the octagon in 
Figure~\ref{octagon} can be written as a conjunction of:
\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
The complementation of the above octagon can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-10-09 introduction.tex
acdcl algorithm

\section{Introduction}
%
A standard static analyses~\cite{se2011} for program safety aims to compute 
invariants by applying transformers that refines an abstract 
value.  Most lattice structure used in static analysis have 
meet and join operations but lack negations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analyzers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behavior.

On the other hand, the combination of precise conjunction in partial assignments 
domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
solvers to reason about disjunction without enumerating cases.  
Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
a formal mechanism for applying abstract interpretation~\cite{se2011} to solve the 
satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
in terms of deduction and abduction transformers, fixed-point, interpolation 
and extrapolation operators on lattices.  The partial assignments in the 
CDCL algorithm are an abstract domain, unit rule is the best abstract 
transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
over the abstract domain, decision is dual widening and clause learning can be 
viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
lifts learning techniques in SAT solver to operate on non-distributive 
lattice -- thus equipping the analysis with negation. This enables the 
analyzer to refine the analysis and prevent enumeration behavior.

The instantiation of ACDCL has so far been limited to interval domain only.  In this paper, 
we identify specific properties of the CDCL algorithm that are necessary to 
lift the propositional skeleton of CDCL to other richer lattice structures 
and relational abstract domains such as octagons, polyhedra, zones and equalities.  
To this end, we present a characterization of abstract model search procedure and 
conflict analysis procedure that lift the abstract deduction transformer and the 
first-UIP algorithm to relational lattice structures.

>>>>

\section{Introduction}
%
Static program analysis based on abstract 
interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
verify certain classes of properties for safety-critical systems.  A 
standard static analyses~\cite{se2011} for program safety aims to compute 
invariants by applying transformers that refines an abstract 
value.  Most lattice structure used in static analysis have 
meet and join operations but lack negations.  The meet operation 
precisely model conjunction, and the join operation over-approximate 
disjunction. Thus, the precision loss in static analyzers is often 
due to imprecise join operations. The loss in precision is overcome 
by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration 
behavior.

On the other hand, the combination of precise conjunction in partial assignments 
domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
solvers to reason about disjunction without enumerating cases.  
Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
a formal recipe for applying abstract interpretation~\cite{se2011} to solve the 
satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
in terms of deduction and abduction transformers, fixed-point, interpolation 
and extrapolation operators on lattices.  The partial assignments in the 
CDCL algorithm are an abstract domain, unit rule is the best abstract 
transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
over the abstract domain, decision is dual widening and clause learning can be 
viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
lifts learning techniques in SAT solver to operate on non-distributive 
lattice -- thus equipping the analysis with negation. This enables the 
analyzer to refine the analysis and prevent enumeration behavior.

The instantiation of ACDCL has so far been limited to interval domains only.  
In this paper, we identify specific properties of the CDCL algorithm that 
are necessary to lift the propositional skeleton of CDCL to other richer 
lattice structures and relational abstract domains such as octagons, 
polyhedra, zones and equalities.  To this end, we present a 
characterization of abstract model search procedure and conflict 
analysis procedure that lift the abstract deduction transformer and the 
first-UIP algorithm to relational lattice structures.
